### 2018-02-15 submission/manuscript.Rmd
update methods section

***Study Analysis Overview:***
OTU richness, evenness, and Shannon diversity were first assessed for differences between controls, adenoma tumors, and carcinoma turmors using both linear mixed-effect models and ORs. For each individual study the Bray-Curtis index was used to assess differences between control-adenoma and control-carcinoma individuals. Next, all common genera were assessed for differences in ORs for individuals having an adenoma or carcinoma and ranked based on P-value. We then built Random Forest models based on the full or selected community (the top 5 increased and top 5 decreased ORs based on P-value). Comparison between the full and selected models took two different approaches. In the first approach, models were trained on one study then tested on the remaining studies using genera-based relative abundances. This process was repeated for every study in the meta-analysis. In the second approach, models were built using OTU level data and a 10-fold CV over 100 different iterations, based on random 80/20 splitting of the data, was used to generate a range of expected AUCs. For these OTU-based models, the selected model included all OTUs that had a taxonomic classification to a taxa in the top 5 increased and top 5 decreased OR analysis based on P-value. Finally, the power of each study was assessed for an effect size ranging from 1% to 30% and an estimated sample size, for these effect sizes, was generated based on 80% power. For comparisons in which only control versus adenoma individuals were made, the carcinoma individuals were excluded from each respective study. Similarly, for comparisons in which control versus carcinoma individuals were made the adenoma individuals were excluded from each respective study. For all analysis completed fecal and tissue samples were kept separate. Within the tissue groups the data were further divided between samples from the same individual (matched) and those from different individuals (unmatched).

***Obtaining Genera Relative Abundance and Selected Models:***
For the genera analysis of the ORs, OTUs were added together based on the genus or lowest available taxonomic classification level and the total average counts, for 100 different subsamplings was obtained. The select models for the first Random Forest model approach utilized the top 5 most significantly increased and decreased ORs. For this approach, the full community models for each study were built by utilizing all genera and lowest taxonomic groups identified within that particular study. The select models for the second approach, that used OTU-based Random Forest models, utilized a similar method as the first approach. The main difference was that any OTU that taxonomically classified to one of the taxa in the top 5 increased or decreased OR analysis were included in the select model. OTU Random Forest models, using the full community, included all OTUs that did not have near zero variance within that specific study.

>>>>

***Study Analysis Overview:***
OTU richness, evenness, and Shannon diversity were first assessed for differences between controls, adenoma tumors, and carcinoma turmors using both linear mixed-effect models and ORs. For each individual study the Bray-Curtis index was used to assess differences between control-adenoma and control-carcinoma individuals. Next, all common genera were assessed for differences in ORs for individuals having an adenoma or carcinoma and corrected for multiple comparisons using the Benjamini-Hochberg method [@benjamini_controlling_1995]. We then built Random Forest models based on all genera, all OTUs, or significant OR taxa (if any were present after multiple comparison correction). For both the full genera and significant OR taxa, models were trained on one study then tested on the remaining studies using genera-based relative abundances. The OTU-based models were built using OTU level data and a 10-fold CV over 100 different iterations, based on random 80/20 splitting of the data, was used to generate a range of expected AUCs. This process was repeated for every study in the meta-analysis. Comparisons of the initial trained model AUCs for the full genera and significant OR taxa were made to the mean AUC generated from the 100 different 10-fold CV runs of the respective OTU-based model. Finally, the power of each study was assessed for an effect size ranging from 1% to 30% and an estimated sample size, for these effect sizes, was generated based on 80% power. For comparisons in which only control versus adenoma individuals were made, the carcinoma individuals were excluded from each respective study. Similarly, for comparisons in which control versus carcinoma individuals were made the adenoma individuals were excluded from each respective study. For all analysis completed fecal and tissue samples were kept separate. Within the tissue groups the data were further divided between samples from the same individual (matched) and those from different individuals (unmatched).

***Obtaining Genera Relative Abundance and Significant OR Taxa Models:***
For the genera analysis of the ORs, OTUs were added together based on the genus or lowest available taxonomic classification level and the total average counts, for 100 different subsamplings was obtained. The significant OR taxa models for the  Random Forest models utilized all taxa that had significant ORs after multiple comparison correction. This meant only models for the carcinoma stool (`r length(crc_OR_stool$measure)` variables) and carcinoma unmatched (`r length(crc_unmatched_OR_tissue$measure)` variables) samples were possible to be created and analyzed.

### 2018-02-15 submission/manuscript.tex
update methods section

\textbf{\emph{Study Analysis Overview:}} OTU richness, evenness, and
Shannon diversity were first assessed for differences between controls,
adenoma tumors, and carcinoma turmors using both linear mixed-effect
models and ORs. For each individual study the Bray-Curtis index was used
to assess differences between control-adenoma and control-carcinoma
individuals. Next, all common genera were assessed for differences in
ORs for individuals having an adenoma or carcinoma and ranked based on
P-value. We then built Random Forest models based on the full or
selected community (the top 5 increased and top 5 decreased ORs based on
P-value). Comparison between the full and selected models took two
different approaches. In the first approach, models were trained on one
study then tested on the remaining studies using genera-based relative
abundances. This process was repeated for every study in the
meta-analysis. In the second approach, models were built using OTU level
data and a 10-fold CV over 100 different iterations, based on random
80/20 splitting of the data, was used to generate a range of expected
AUCs. For these OTU-based models, the selected model included all OTUs
that had a taxonomic classification to a taxa in the top 5 increased and
top 5 decreased OR analysis based on P-value. Finally, the power of each
study was assessed for an effect size ranging from 1\% to 30\% and an
estimated sample size, for these effect sizes, was generated based on
80\% power. For comparisons in which only control versus adenoma
individuals were made, the carcinoma individuals were excluded from each
respective study. Similarly, for comparisons in which control versus
carcinoma individuals were made the adenoma individuals were excluded
from each respective study. For all analysis completed fecal and tissue
samples were kept separate. Within the tissue groups the data were
further divided between samples from the same individual (matched) and
those from different individuals (unmatched).

\textbf{\emph{Obtaining Genera Relative Abundance and Selected Models:}}
For the genera analysis of the ORs, OTUs were added together based on
the genus or lowest available taxonomic classification level and the
total average counts, for 100 different subsamplings was obtained. The
select models for the first Random Forest model approach utilized the
top 5 most significantly increased and decreased ORs. For this approach,
the full community models for each study were built by utilizing all
genera and lowest taxonomic groups identified within that particular
study. The select models for the second approach, that used OTU-based
Random Forest models, utilized a similar method as the first approach.
The main difference was that any OTU that taxonomically classified to
one of the taxa in the top 5 increased or decreased OR analysis were
included in the select model. OTU Random Forest models, using the full
community, included all OTUs that did not have near zero variance within
that specific study.

@@

>>>>

\textbf{\emph{Study Analysis Overview:}} OTU richness, evenness, and
Shannon diversity were first assessed for differences between controls,
adenoma tumors, and carcinoma turmors using both linear mixed-effect
models and ORs. For each individual study the Bray-Curtis index was used
to assess differences between control-adenoma and control-carcinoma
individuals. Next, all common genera were assessed for differences in
ORs for individuals having an adenoma or carcinoma and corrected for
multiple comparisons using the Benjamini-Hochberg method {[}51{]}. We
then built Random Forest models based on all genera, all OTUs, or
significant OR taxa (if any were present after multiple comparison
correction). For both the full genera and significant OR taxa, models
were trained on one study then tested on the remaining studies using
genera-based relative abundances. The OTU-based models were built using
OTU level data and a 10-fold CV over 100 different iterations, based on
random 80/20 splitting of the data, was used to generate a range of
expected AUCs. This process was repeated for every study in the
meta-analysis. Comparisons of the initial trained model AUCs for the
full genera and significant OR taxa were made to the mean AUC generated
from the 100 different 10-fold CV runs of the respective OTU-based
model. Finally, the power of each study was assessed for an effect size
ranging from 1\% to 30\% and an estimated sample size, for these effect
sizes, was generated based on 80\% power. For comparisons in which only
control versus adenoma individuals were made, the carcinoma individuals
were excluded from each respective study. Similarly, for comparisons in
which control versus carcinoma individuals were made the adenoma
individuals were excluded from each respective study. For all analysis
completed fecal and tissue samples were kept separate. Within the tissue
groups the data were further divided between samples from the same
individual (matched) and those from different individuals (unmatched).

\textbf{\emph{Obtaining Genera Relative Abundance and Significant OR
Taxa Models:}} For the genera analysis of the ORs, OTUs were added
together based on the genus or lowest available taxonomic classification
level and the total average counts, for 100 different subsamplings was
obtained. The significant OR taxa models for the Random Forest models
utilized all taxa that had significant ORs after multiple comparison
correction. This meant only models for the carcinoma stool (8 variables)
and carcinoma unmatched (3 variables) samples were possible to be
created and analyzed.

@@

\hypertarget{ref-benjamini_controlling_1995}{}
51. Benjamini Y, Hochberg Y. Controlling the false discovery rate: A
practical and powerful approach to multiple testing. Journal of the
Royal Statistical Society Series B (Methodological). 1995;57:289--300.
